Topological Equivariant Artist Model for Visualization Library Architecture

### 2025-03-13 paper/main.tex
small clean ups/restors

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
The abstract goes here.
\end{abstract}

@@

\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
A \textbf{functor} must satisfy the properties
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
\end{itemize}
$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to.
\end{definition}
Modeling the data container as a functor allows us state that, just like a functor, the container is a map between index space objects and sets of data records that preserve morphisms between index space objects and data records.
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
\begin{definition}\label{def:atct:presheaf}
  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}.
\end{definition}
A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects. The presheaf is contravariant because the inclusion morphisms between input object $\iota: \opensetc_1 \rightarrow \opensetc_2$
are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections $\iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$ restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.

@@

\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \imartist{\gbasec}{\gtotalc}, \imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}
\end{align}

@@

While Artists are defined as maps going from data over \dbase\ to graphic over \gbase\, they are compatible with maps defined solely over \dbase\ or \gbase\. This is because when a sheaf is equipped with transport functors (\vindexpush, \vindexpull), maps between sheaves over one space $Hom _{\sheaf_{\dbase}}$ are isomorphic to maps between sheaves over the other space $Hom _{\sheaf_{\gbase}}$\cite{harder2008lectures}:

@@

\subsubsection{Addition}
\label{sec:artist:addition}
We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
  The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from. Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code. We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc. - in a manner that independently encodes the mathematical structure of the topology and the fields of the dataset. Modeling the data structures that store the datasets as sheaves provide a method for specifying visualization methods that are designed to work regardless of how the dataset is stored - whether the data is on disk, distributed, or on demand. Specifying the visualization library components as natural transforms of sheaves means that the constraints that the component must satisfy to be structure preserving can be specified as the set of morphisms on the data and graphic sheaves, including the structure on the topology and fields of the data.
  Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
\end{abstract}

@@

\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
A \textbf{functor} must satisfy the properties
\begin{itemize}
  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
\end{itemize}
$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to.
\end{definition}
Modeling the data container as a functor allows us state that, just like a functor, the container is a map between index space objects and sets of data records that preserve morphisms between index space objects and data records.
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
\begin{definition}\label{def:atct:presheaf}
  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{spanier1989algebraic}.
\end{definition}

A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects. The presheaf is contravariant when for every arbitrary morphism between input base spaces $\dfunch: \openset_1 \rightarrow \openset_2$ there exists a corresponding pullback function between the sets of sections $\dfuncpull: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$.

%% these two paragraphs cen be condensed and moved below the figure as an example
A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects. The presheaf is contravariant because the inclusion morphisms between input object $\iota: \opensetc_1 \rightarrow \opensetc_2$
are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections $\iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$ restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.

@@

\begin{equation*}
  \begin{tikzcd}
    & {} \arrow[dd, "\texttt{artist}" description, Rightarrow, shift left=3.65, shorten <= .7em, shorten >= .7em, color=artist, font=\Huge]
    & \\ {\scriptstyle\textcolor{base}{\texttt{topology}}}
      \arrow[rr, "\texttt{graphic}"', bend right, color=sheaf, font=\huge] \arrow[rr, "\texttt{data}", bend left, color=sheaf, font=\huge] &
    & {\scriptstyle\textcolor{base}{\texttt{topology}}\textcolor{section}{\rightarrow} \textcolor{fiber}{\texttt{fields}}} \\
    & {}                                              &
\end{tikzcd}
\end{equation*}

%move this down to what is math section
\begin{align}
  \label{eq:artist:hom_transport}
  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \imartist{\gbasec}{\gtotalc}, \imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}
\end{align}

@@

While Artists are defined as maps going from data over \dbase\ to graphic over \gbase\, they are compatible with maps defined solely over \dbase\ or \gbase. This is because when a sheaf is equipped with transport functors (\vindexpush, \vindexpull), maps between sheaves over one space $Hom _{\sheaf_{\dbase}}$ are isomorphic to maps between sheaves over the other space $Hom _{\sheaf_{\gbase}}$\cite{harder2008lectures}:

@@

\subsubsection{Addition}
\label{sec:artist:addition}
We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset
\note{adjoint equation is wrong, probably multiplication b/c using $k^a$ and $K^b$ to define the piecewise, disjoint union is coproduct}
